---
title: "Meta Analysis_Two Approaches"
output: github_document
---

##AIM
To explain Meta analytic approach in random effects model (REM) framework using 1. summary statistics (SSM) and 2. Fully Bayesian (FBM) method. 


**Ref: Meta-analysis: formulating, evaluating, combining, and reporting Normand,ST Statist. Med. 1999**

REM here indicates combining multiple (k) studies each with a study effect $y_{i}$ where $i = 1,2,3.....k$ 

##AIM 1 - SSM:


Underlying model is <br><br>$y_{i} \sim N(\mu_{i},v^{2}_{i})$ <br><br> $\mu_{i} \sim N(d,\sigma_{\mu}^{2})$. <br><br>Assumption is $v^{2}_{i}$ are known (estimated from data, most of the cases asymptotic variance) <br><br>quantities of interest are  $\mu$ and $\sigma_{\mu}^{2}$.

##Step 1:Summary Statistics

We use metafor pacakge to estimate the summary measure $y_{i}$ and asymptotic variance $v^{2}_{i}$

In this note $y_{i}$ is log odds ratio of study i $\forall 1\le i \le k$ and hence $v^{2}_{i} = a^{-1}+b^{-1}+c^{-1}+d^{-1}$ where $a,b,c,d$ are cell counts of given $2\times2$ table

```{r echo = TRUE, warning = FALSE,message=FALSE}
#Summary Statistics Approach
#Step 1
ami=c(0,	0,	0,	0,	1,	0,	0,	2,	0,	0,	0,	0,	0,	1,	0,	0,	1,	1,	0,	0,	0)
bmi=c(100,	100,	28,	14,	19,	80,	54,	148,	60,	15,	10,	15,	88,	141,	136,	204,	97,	23,	25,	41,	21)
cmi=c(0,	0,	1,	0,	1,	0,	0,	0,	2,	0,	0,	1,	2,	2,	0,	2,	2,	1,	0,	0,	0)
dmi=c(100,	101,	36,	20,	18,	80,	49,	150,	58,	16,	10,	19,	86,	137,	131,	182,	97,	25,	25,	67,	16)
require(metafor)
y=escalc(ai=ami,bi=bmi,ci=cmi,di=dmi,measure = "OR",add = 10^(-8), to="only0" ,  drop00=FALSE)
#End of step 1----------------------------------------------------------
```


#Step 2: Stan code for REM Normal-Normal Model 
```{r engine='stan',engine.opts=list(x = "MA_SS_tbyt")}
data {
  int<lower=0> N; 
  real y[N]; 
  real<lower=0> v1[N]; 
} 
parameters {
  real d; 
  real<lower=0> sigma_mu; 
  vector[N] mu;
} 
transformed parameters {
  real<lower=0> sigmasq_mu; 
  sigmasq_mu=sigma_mu*sigma_mu;
} 
model {
  y ~ normal(mu, v1);
  mu ~ normal(d, sigma_mu);
  d ~ normal(0, 1E3); 
  sigmasq_mu ~ inv_gamma(1E-3, 1E-3); 
}
```


```{r}
require(rstan)
yi=as.vector(y$yi);v1=sqrt(y$vi) #sq.root is for SD parameter in Normal in Stan
N=length(ami)
MA_SS_data <- list(y=yi,v=v1,N=N)
MA_SS_init <- function(){list(
  d = 0,
  mu = rnorm(N,0,1),
  sigma_mu = 1
)}
mod1 <- sampling(MA_SS_tbyt,data=MA_SS_data, init = MA_SS_init,
                 control = list(adapt_delta=0.99,max_treedepth=15),iter=2000,chains=4) 
fitNor_Nor_summary<-as.data.frame(summary(mod1, pars = c("d","sigmasq_mu"), probs = c(0.025, 0.975))$summary)
ss_re=round(fitNor_Nor_summary,4)
```




##AIM 2 - FBM:
**Additional Reference:Smith et al 1995 Bayesian Approaches to Random-Effects Meta analysis: A Comparative Study**


Underlying model is  
$r^{C}_{i} \sim B(p^{C}_{i},n^{C}_{i})$

$r^{T}_{i} \sim B(p^{T}_{i},n^{T}_{i})$

logit($p^{C}_{i}$)=$\mu_{i}$-$\delta_{i}/2$

logit($p^{T}_{i}$)=$\mu_{i}+\delta_{i}/2$

$\delta_{i}\sim N(d,\sigma^{2})$

$\mu_{i}\sim N(d_{0},\sigma^{2}_{0})$


Here, $r^{C}_{i}$ indicates the number of successes in Control group arising from  $n^{C}_{i}$ cases assumed to have probability of $p^{C}_{i}$ 
<br><br>Similarly $r^{T}_{i}$ can be defined for treatment group
<br><br>Quantities of interest are $d$ and $\sigma^{2}$.


#Fully Bayesian Method

**Stan Model** 

```{r engine='stan',engine.opts=list(x = "MA_FB_tbyt")}
data {
  int<lower=0> N; 
  int<lower=0> nt[N]; 
  int<lower=0> rt[N]; 
  int<lower=0> nc[N]; 
  int<lower=0> rc[N]; 
} 
parameters {
  real d; 
  real<lower=0> sigma_delta; 
  vector[N] mu;
  vector[N] delta;
} 
transformed parameters {
  real<lower=0> sigmasq_delta; 
  sigmasq_delta = sigma_delta*sigma_delta; 
  
} 
model {
  rt ~ binomial_logit(nt, mu+delta/2);
  rc ~ binomial_logit(nc, mu-delta/2);
  delta  ~ normal(d, sigma_delta); 
  mu ~ normal(0, 2);
  d ~ normal(0, sqrt(10)); 
  sigmasq_delta ~ inv_gamma(3, 1); 
}
```




```{r}
require(rstan)
nt=ami+bmi;nc=cmi+dmi
N=length(ami)

MA_FB_data <- list(nc=nc,rc=cmi,nt=nt,rt=ami,N=N)
MA_FB_init <- function(){list(
  d = 0,
  mu = rnorm(N,0,1),
  delta= rnorm(N,0,1),
  sigma_delta = 1
)}
fitMA_FB <- stan(file="MA_FB_tbyt.stan",
                 data=MA_FB_data, init = MA_FB_init,
                 control = list(adapt_delta=0.99,max_treedepth=15),iter=2000,chains=4)

fitMA_FB_summary<-as.data.frame(summary(fitMA_FB, pars = c("d","sigmasq_delta"), probs = c(0.025, 0.975))$summary)
fb_re=round(fitMA_FB_summary,4)
```

```{r message=FALSE}
library(knitr)
kable(ss_re,caption = "Summary Statistics Method")
kable(fb_re,caption = "Fully Bayesian Method")
```



This demonstrates the two MA approaches. Summary statistics requires continuity corrections for zero cells, which requires a careful investigation; where as this is completely alleviated when FBM is used.

Still a careful specification of priors (especially for between variance) is always a concern in Bayesian analysis